---
title: "Fifa_Overall"
author: "Adam Johnson"
date: "12/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Linear Regression to Predict FIFA Player Overalls

In this code I will be building a linear model to attempt to player rankings from the video game Fifa.  The data set used is provided here: https://www.kaggle.com/karangadiya/fifa19

```{r}
# Loading the Data and Removing any NAs
fifa <- read.csv("fifa_data.csv")
fifa <- na.omit(fifa)
```

## Overview of the data

```{r}
# Summary Stats of all of the Columns
head(fifa)
summary(fifa)

```

As I will be attempting to predict Overall, I created the below histogram to get a better sense of how the data is distributed. 

```{r}
# Creates histogram of the Dependent Variable 
hist(fifa$Overall)
```

It appears that the variable that I will be trying to predict is normally distributed. 


Next, I went in and selected only the Numeric Columns that I will be using as predictors.  I also removed ID columns and columns related to Goal Keeping stats, as they are irrelevant to the majority of players. 
```{r}
# Selecting Just the Numeric Columns
num_column <- unlist(lapply(fifa, is.numeric))
num_fifa <- fifa[ ,num_column]

# Also removing Identification Columns along with columns related to GK
num_fifa <- subset(num_fifa, select = -c(ID, ï.., GKPositioning,GKDiving,GKReflexes, GKHandling, GKKicking))

head(num_fifa)
```

## Visualizations

I used the code below to make scatter plots of all of the variables put against Overall ranking. This can help to visualize which of the relationships are linear.  
```{r}
# Generates scatter plots of all of the variables against Overall
library("dplyr")
library("ggplot2")
library("tidyverse")
num_fifa %>%
  gather(-Overall, key = "var", value = "value") %>%
  ggplot(aes(x = value, y = Overall)) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~ var, scales = "free") +
  theme_bw()
```

Next, I wanted to check for correlations between the variables.  This can help to get an idea of which predictors to use in the model. 
```{r}
# Code to visually check for correlations
library(corrplot)
correlations <- cor(num_fifa)
corrplot(correlations, method="circle")
```
As there appear to be many correlations between the different variables, I will start the modeling with all of the variables included, and adjust from there.  

## Partitioning Data 

Before training the model, I will create two data sets.  One for training and one for testing the resulting model.  

```{r}
# Creates training and test data
set.seed(100)  # setting seed to reproduce results of random sampling
trainingRow <- sample(1:nrow(num_fifa), 0.8*nrow(num_fifa))  
trainingData <- num_fifa[trainingRow, ]  
testData  <- num_fifa[-trainingRow, ]  
```

## Modeling 

For the first attempt at the model, I went in and used all of the available predictors. This is the base line from which the model can be adjusted in future iterations. 
```{r}
# Creates the first model. 
fifa_model <- lm(Overall ~ ., data = trainingData)
# Creates summary of the model to review the metrics. 
summary(fifa_model)
# Creates plots of the model to be used for validation purposes. 
par(mfrow = c(2,2))
plot(fifa_model)
```


As many of the variables were not significant, for the next model, I went ahead and removed all of them,  

```{r}
# Removing Non-significant Coefficients to compare to previous model.  
trainingData <- subset(trainingData, select = -c(Weak.Foot, Curve, FKAccuracy, Agility, LongShots, Interceptions, Vision, Marking, StandingTackle))
                                        
```

Now that all of the non-significant coefficients have been removed, I once again ran the linear model.

```{r}
# Second rendition of the model, this time will all of the non-significant coefficients removed.
fifa_model2 <- lm(Overall ~ ., data = trainingData)
summary(fifa_model2)
par(mfrow = c(2,2))
plot(fifa_model2)
```

New model has improved F-Statistics, while all other attributes are the same. All coefficients are now significant. 

To see if the model can be improved any more, I also removed the two coefficients with the lowest levels of significance.  As we can see from the previous model summary, these two coefficients are Finishing and Volleys. 

```{r}
# Removing lowest significance to see if any improvement
trainingData <- subset(trainingData, select = -c(Finishing, Volleys))
```

Now that all of the selected variables have been removed, it is time to once again run the model.  

```{r}
# Third rendition of the model. 
fifa_model3 <- lm(Overall ~ ., data = trainingData)
summary(fifa_model3)
par(mfrow = c(2,2))
plot(fifa_model3)
```
Third model results in all highly significant coefficients.  Highest F-Statistic so far, however the R^2 value dropped by a minuscule 0.01.  I will be using this third model for making my predictions as it has the highest F-statistic, which indicates that it has the best fit of the three models. 


## Predictions

Now that the model has been chosen, it is time to actually make predictions with the model.  

```{r}
# Creates the predictions and saves in predictions variable
predictions <- round(predict(fifa_model3, trainingData),0)

# Creates data frame of predictions and actual values
actuals_preds <- data.frame(cbind(actuals=trainingData$Overall, predicteds=predictions))
head(actuals_preds)

# Takes correlation between the actual values and the predicted values
cor_accuracy <- cor(actuals_preds)
cor_accuracy

# Calculates min max accuracy of the predictions. The closer to 1, the better the model is performing. 
min_max_accuracy <- mean(apply(actuals_preds, 1, min) / apply(actuals_preds, 1, max))  
min_max_accuracy 

```

When running the model on the training data, it is clear that the model is performing very well.  The correlation between the actual values and the predictions is approximately 95%.  This indicates that the predictions are very close to the actual values.  The calculated min max value is .9757, which also indicates that the model is performing very strongly.

Finally, to confirm that this model is working well, I did the same predictions on the test data.  This is to check to see if there is any over fitting, and if the model can still perform well on a data set that it was not trained on.  

```{r}
# Running Model on Testing data to see if there is any over fitting

predictions <- round(predict(fifa_model3, testData),0)

actuals_preds <- data.frame(cbind(actuals=testData$Overall, predicteds=predictions))
head(actuals_preds)
cor_accuracy <- cor(actuals_preds)
cor_accuracy

min_max_accuracy <- mean(apply(actuals_preds, 1, min) / apply(actuals_preds, 1, max))  
min_max_accuracy  


```

The predictions on the test data resulted in nearly identical correlations and min max values.  Because of this, we can say that there does not appear to be a risk of over fitting.  






















